MSstatsPTM statistical relative quantification of post-translational modifications in global proteomics experiments \textbf{Supplementary Information}

### 2022-03-05 supplementary/ptm_sm.tex
experimental design changes

\title{Statistical methods for relative quantification of post-translational modifications in global proteomics experiments}
\author[1]{Tsung-Heng~Tsai}
\author[2]{Erik~Verschueren}
\author[1]{Olga~Vitek}
\affil[1]{Northeastern University, Boston, MA, USA}
\affil[2]{Protein Chemistry, Genentech, South San Francisco, CA, USA}

\begin{document}
%\maketitle

%\newpage
\tableofcontents

@@

This study addresses three major goals in the characterization of post-translational modifications (PTMs): a) relative PTM quantification, b) PTM significance analysis, i.e., to detect PTM sites that are differentially modified across experimental conditions, and c) statistical design of PTM experiments.

@@

The proposed statistical approach was evaluated and compared to the $t$-test and Limma methods using computer simulation. Specifically, their statistical properties with respect to protein-level adjustment and real world experimental conditions were evaluated.

@@

Again we consider three different methods and assess their performance: the method proposed in this paper, Limma, and two sample $t$-test. All methods are analyzed after adjusting for changes in overall protein level. The proposed method summarizes feature intensities up to the run level using Tukey's Median Polish, while the other methods use the log of summed feature intensities. The results are summarized from Figure \ref{fig:spike_volcano_msstats} to Figure \ref{fig:spike_fc}, including volcano plots, model summary statistics, and fold change analysis.

>>>>

\title{MSstatsPTM statistical relative quantification of post-translational modifications in global proteomics experiments

\textbf{Supplementary Information}
}

\author[1]{Devon~Kohler}
\author[2]{Tsung-Heng~Tsai}
\author[1]{Ting~Huang}
\author[4]{Erik~Verschueren}
\author[3]{Trent~Hinkle}
\author[3]{Meena~Choi*}
\author[1]{Olga~Vitek*}
\affil[1]{Khoury College of Computer Science, Northeastern University, Boston, MA, USA}
\affil[2]{Kent State University, Kent, OH, USA}
\affil[3]{MPL, Genentech, South San Francisco, CA, USA}
\affil[4]{ULUA BV, Arendstraat 29, 2018 Antwerp, Belgium}
\affil[*]{Corresponding Authors}

\begin{document}
\maketitle

\newpage
\tableofcontents

@@

This study addresses three major goals in the characterization of post-translational modifications (PTMs): a) relative PTM quantification, b) PTM significance analysis, i.e., to detect PTM sites that are differentially modified across experimental conditions, and c) statistical design and analysis of PTM experiments.

@@

The proposed statistical approach was evaluated and compared to the $t$-test and Limma methods using computer simulations. Specifically, their statistical properties with respect to protein-level adjustment and real world experimental conditions were evaluated.

@@

Again we consider three different methods and assess their performance: the method proposed in this paper, Limma, and two sample $t$-test. All methods are analyzed after adjusting for changes in overall protein level. The proposed method summarizes feature intensities up to the run level using Tukey's Median Polish, while the other methods use the log of summed feature intensities. The results are summarized from Figure \ref{fig:spike_volcano_msstats} to Figure \ref{fig:spike_stat}, including volcano plots, model summary statistics, and fold change analysis.
